Geometric variational problems

### 2020-11-16 NSF2020/Description/specificaim1_ver2.tex

%  Past researchers expressed the strain tensor in terms of monolayer tangent plane.   
%  In this frame, however, there is a coupling between the gradients of the monolayer surface, lipid direction, and chain length,
%  and accounting for these couplings by Taylor expansion  
%  can lead to an energy density with an unruly number of parameters \cite{PhysRevE.102.042406}.
%
%The form of the elastic energy density \eqref{ansatz3} is the same as
%  the Oseen-Frank energy density for nematic liquid crystals \cite{ANDRIENKO2018520,Tran7106,Helfrich73}.   In fact,  
%  a lipid monolayer acts as one layer in a smectic  phase \cite{REYESMATEO1995978,Rangamani20140463,PhysRevLett.113.248102}. 
%  Motivated by liquid crystal theory, we have found that the gradient terms completely decouples when 
%  expanding the strain tensor, not in terms of the monolayer tangent plane, but in terms of a plane perpendicular to ${\bf n}$. 
%  Using this observation, we prove an exact identity 
%  that gives the incompressibility condition by a Steiner-type polynomial in $\Div {\bf n}$ and $\det \nabla {\bf n}$ \cite{Fe59}.
%  In contrast, the works \cite{TerziDeserno17, PhysRevE.102.042406, Hamm2000, C9SM02079A} utilize an
%  approximate identity for incompressibility, and so we can offer a considerably more accurate analysis of monolayer energy.
%  
%%To measure the twist modulus, we place a single monolayer on a hydrophobic substrate 
%%and take the initial configuration ${\bf n} = (0, f(x_1), \sqrt{1 - f^2(x_1)})$,
%%yielding a surface gradient $\nabla {\bf n} = \begin{pmatrix}0 & 0\\ f'(x_1) & 0\end{pmatrix}$. 
%%  Here $\Div {\bf n} = \mathrm{tr}(\nabla {\bf n}) = 0$, $\det(\nabla {\bf n}) = 0$ and
%%  $\Curl {\bf n} = (\nabla u)_{12} - (\nabla u)_{21} = f'(x_1)$.
%%  This gives the continuum elastic energy equals
%%  \begin{equation}
%%    \label{eq:twist_energy}
%%    E = \int_{\Sigma} \tfrac{1}{2}\KT (f'(x_1))^2 + \tfrac{1}{2} \KTH f^2(x_1) \,\dif dx_1 \dif x_2,
%%  \end{equation}
%%  which we compare to the HAP simulation data.
%
\subsubsection{Existence of minimizers}
The HK functional \eqref{ansatz3} has had tremendous impact in understanding
biological function, but has received little attention in terms of mathematical
analysis. In the calculus of variations, the principal question is whether a
minimizer of an energy functional exists. But in the case of \eqref{ansatz3},
this answer is presently unknown. With regard to the simpler, Canham-Helfrich
energy, \cite{Simon1993} proved the existence of minimizing surfaces without boundary, 
and \cite{doi:10.1137/18M1195851} proved the well-posedness of a spatially periodic,
time-dependent elastic interface problem. An analytical challenge 
for the HK functional \eqref{ansatz3} is that coupling to a director field makes it
possible to have bounded energy monolayers with corners, and such
pathological examples must be ruled out before carrying over the arguments
for the Canham-Helfrich functional to the present setting.

\subsubsection{Asymptotic analysis on the limit $\rho\rightarrow 0$ of HAP}
Our long term goal is to develop tools that can explain functionals like
\eqref{ansatz3} from first principles in terms of HAP. The papers
\cite{doi:10.1063/5.0009734,Seguin2012,Seguin2014} give statistical
mechanical/mean field derivations of the Canham-Helfrich energy from a
pair-potential for rod-like molecules, but do not include tilt, which is an
indispensable deformation at biological scales. To make progress, we must
first understand how the HAP functional behaves under various limits.

>>>>

%  Past researchers expressed the strain tensor in terms of monolayer tangent plane.   
%  In this frame, however, there is a coupling between the gradients of the monolayer surface, lipid direction, and chain length,
%  and accounting for these couplings by Taylor expansion  
%  can lead to an energy density with an unruly number of parameters \cite{PhysRevE.102.042406}.
%
%The form of the elastic energy density \eqref{ansatz3} is the same as
%  the Oseen-Frank energy density for nematic liquid crystals \cite{ANDRIENKO2018520,Tran7106,Helfrich73}.   In fact,  
%  a lipid monolayer acts as one layer in a smectic  phase \cite{REYESMATEO1995978,Rangamani20140463,PhysRevLett.113.248102}. 
%  Motivated by liquid crystal theory, we have found that the gradient terms completely decouples when 
%  expanding the strain tensor, not in terms of the monolayer tangent plane, but in terms of a plane perpendicular to ${\bf n}$. 
%  Using this observation, we prove an exact identity 
%  that gives the incompressibility condition by a Steiner-type polynomial in $\Div {\bf n}$ and $\det \nabla {\bf n}$ \cite{Fe59}.
%  In contrast, the works \cite{TerziDeserno17, PhysRevE.102.042406, Hamm2000, C9SM02079A} utilize an
%  approximate identity for incompressibility, and so we can offer a considerably more accurate analysis of monolayer energy.
%  
%%To measure the twist modulus, we place a single monolayer on a hydrophobic substrate 
%%and take the initial configuration ${\bf n} = (0, f(x_1), \sqrt{1 - f^2(x_1)})$,
%%yielding a surface gradient $\nabla {\bf n} = \begin{pmatrix}0 & 0\\ f'(x_1) & 0\end{pmatrix}$. 
%%  Here $\Div {\bf n} = \mathrm{tr}(\nabla {\bf n}) = 0$, $\det(\nabla {\bf n}) = 0$ and
%%  $\Curl {\bf n} = (\nabla u)_{12} - (\nabla u)_{21} = f'(x_1)$.
%%  This gives the continuum elastic energy equals
%%  \begin{equation}
%%    \label{eq:twist_energy}
%%    E = \int_{\Sigma} \tfrac{1}{2}\KT (f'(x_1))^2 + \tfrac{1}{2} \KTH f^2(x_1) \,\dif dx_1 \dif x_2,
%%  \end{equation}
%%  which we compare to the HAP simulation data.
%
\subsubsection{Analysis on the HAP model in terms of the HK functional}
The HK functional \eqref{ansatz3} has had tremendous impact in understanding
biological membranes, but has received little attention in terms of mathematical
analysis. In the calculus of variations, the principal question is whether a
minimizer of an energy functional exists. But in the case of \eqref{ansatz3},
this answer is presently unknown. With regard to the simpler, Canham-Helfrich
energy, \cite{Simon1993} proved the existence of minimizing surfaces without boundary, 
and \cite{doi:10.1137/18M1195851} proved the well-posedness of a spatially periodic,
time-dependent elastic interface problem. An analytical challenge 
for the HK functional \eqref{ansatz3} is that coupling to a director field makes it
possible to have bounded energy monolayers with corners, and such
pathological examples must be ruled out before carrying over the arguments
for the Canham-Helfrich functional to the present setting.

%\subsubsection{Asymptotic analysis on the limit $\rho\rightarrow 0$ of HAP}
Our goal here is to develop tools that can explain functionals like
\eqref{ansatz3} from first principles in terms of HAP. The papers
\cite{doi:10.1063/5.0009734,Seguin2012,Seguin2014} give statistical
mechanical/mean field derivations of the Canham-Helfrich energy from a
pair-potential for rod-like molecules, but do not include tilt, which is an
indispensable deformation at biological scales. To make progress, we must
first understand how the HAP functional behaves under various limits.
